---
title: "Investigating the Use of Generalised Additive Models"
author: "Mick Cooney <mickcooney@gmail.com>"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: spacelab #sandstone #spacelab #flatly
    highlight: pygments
    number_sections: TRUE
    toc: TRUE
    toc_depth: 2
    toc_float:
      smooth_scroll: FALSE
  pdf_document: default
---


```{r knit_opts, include=FALSE}
library(conflicted)
library(tidyverse)
library(magrittr)
library(rlang)
library(scales)
library(purrr)
library(forcats)
library(vctrs)
library(cowplot)
library(mgcv)
library(broom)
library(CASdatasets)


#source("custom_functions.R")

conflict_prefer("filter",          "dplyr")
conflict_prefer("select",          "dplyr")
conflict_prefer("lag",             "dplyr")



knitr::opts_chunk$set(tidy       = FALSE,
                      cache      = FALSE,
                      message    = FALSE,
                      warning    = FALSE,
                      fig.height =     8,
                      fig.width  =    11)



options(width = 80L,
        warn  = 1,
        mc.cores = parallel::detectCores()
        )

theme_set(theme_cowplot())


stan_seed <- 42
set.seed(42)
```


# Load `rock` Data

First we need to load the various datasets we are using in our model.

```{r load_datasets, echo=TRUE}
data(rock)

rock %>% glimpse()
```

In our models, we are trying to predict the value for permeability, `perm`
based on the other measures.


## Scatterplots for Variables

We want to build quick scatterplots for each of the three input variables
against `perm`.

```{r construct_scatterplots, echo=TRUE}
plot_tbl <- rock %>%
  gather('input', 'value', -perm)

ggplot(plot_tbl) +
  geom_point(aes(x = value, y = perm)) +
  facet_wrap(vars(input), scales = "free", ncol = 2) +
  xlab("Input Value") +
  ylab("Permeability")
```

We also want to repeat this plot against the log of the permeability.

```{r}
plot_tbl <- rock %>%
  gather('input', 'value', -perm) %>%
  mutate(log_perm = log(perm))

ggplot(plot_tbl) +
  geom_point(aes(x = value, y = log_perm)) +
  facet_wrap(vars(input), scales = "free", ncol = 2) +
  xlab("Input Value") +
  ylab("Log of Permeability")
```


# Fit Regression Models

We start by fitting a basic linear model first, fitting to `log_perm`.


```{r fit_linear_model, echo=TRUE}
rock_lm <- lm(log(perm) ~ area + peri + shape, data = rock)

rock_lm %>% glance()
```



```{r fit_gam_model, echo=TRUE}
rock_gam <- gam(log(perm) ~ s(area) + s(peri) + s(shape), 
                data = rock,
                control = gam.control(maxit = 50)
                )

rock_gam %>% summary()
rock_gam %>% glance()
```


```{r compare_models_anova, echo=TRUE}
anova(rock_lm, rock_gam)
```

```{r plot_different_models, echo=TRUE}
par(mfrow = c(2, 3), pty = "s")
plot(rock_gam, se = T)

rock_gam1 <- gam(log(perm) ~ area + peri + s(shape), data = rock)
plot(rock_gam1, se = T)


anova(rock_lm, rock_gam1, rock_gam)
```
